---
title: "Andrew Bowen DATA607 Tidyverse Create"
author: "Andrew Bowen"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, message=FALSE}
library(tidyverse)
```

I chose the [House of Representatives historical polling dataset from FiveThirtyEight](https://projects.fivethirtyeight.com/polls/). I'd like to read in the csv file (hosted on our GitHub) and convert it into a `tibble` to demo some [tidyverse functions](https://www.tidyverse.org/packages/) available to [`tibble` objects](https://tibble.tidyverse.org)
```{r}
data_url <- "https://raw.githubusercontent.com/andrewbowen19/FALL2022TIDYVERSE/main/data/house_polls_historical.csv"

df <- read.csv(data_url)

# Converting
tbl_df <- as_tibble(df)
head(tbl_df)
```


Showing off some `forcats` functions as well. Let's see a count of total polls by pollster, using the `fct_count` function on our tibble.
```{r}
fct_count(tbl_df$pollster, sort=TRUE)
```


Let's plot the polling percentages of the major parties (republicans and democrats)
```{r}
major_parties <- tbl_df %>%
                        filter(party == "DEM" | party == "REP")

ggplot(major_parties, aes(x = pct, color=party)) +
    geom_histogram(binwidth=5, fill="white") +
    scale_color_manual(values=c("blue", "red"))

```

Let's also make a bar chart to show the average polling % by party. Let's also make a bar graph showing the parties with an average polling percentage above 3%
```{r}
avg_pct_by_party <- tbl_df %>%
      group_by(party) %>%
      summarise(avg_pct = mean(pct)) %>%
      filter(avg_pct > 3.0)

ggplot(avg_pct_by_party, aes(x = party, y= avg_pct)) + geom_bar(stat='identity') +
    ylab("Average Polling %") + xlab("Political Party")
```

