---
title: "Tidyverse CREATE"
author: "Benjamin Inbar"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<div style="margin-top:50px;">
##### Using one or more TidyVerse packages, and any dataset from fivethirtyeight.com or Kaggle, create a programming sample “vignette” that demonstrates how to use one or more of the capabilities of the selected TidyVerse package with your selected dataset.
First, tidyverse!
```{r warnings=FALSE, message=FALSE}
library(tidyverse)
library(knitr)
```

<div style="margin-top:50px;">
Read in the Donald Trump favorability ratings data from <a href="https://projects.fivethirtyeight.com/polls/">FiveThirtyEight</a>
```{r}
favdata <- read_csv("https://projects.fivethirtyeight.com/polls/data/favorability_polls.csv", show_col_types=FALSE)
head(favdata)
colnames(favdata)
```

<div style="margin-top:50px;">
Some of the most useful tidyverse functions come from dplyr, which we can use to easily select, group, and summarize useful information.
```{r}
fav_unfav <- favdata |> select(favorable, unfavorable)
favdata <- favdata |> mutate(responding=rowSums(fav_unfav))
kable(head(favdata[c("favorable", "unfavorable", "responding")]))
```

<div style="margin-top:50px;">
Get average population size, and average number responding, per pollster.
```{r}
averages <- favdata |> group_by(pollster) |> summarize(Avg_Sample_Size=mean(sample_size), Avg_Responding=mean(responding))
kable(averages)
```

<div style="margin-top:50px;">
Plot the relationship between sample size and response rate. Notable that larger samples give higher response rates. Maybe because those polls have more resources to contact all those respondents?
```{r message=FALSE}
ggplot(data=averages, aes(x=Avg_Sample_Size, y=Avg_Responding)) +
  geom_point(size=2, color='darkblue', shape=16) +
  geom_smooth(method = 'lm', se=FALSE, color='red')
```

<div style="margin-top:50px;">
Nesting is part of the 'purr' tidyverse package and allows for summarizing of rows, simliar to dplyr's group_by().
```{r}
favdata_nested <- favdata |> group_by(pollster) |> nest()
favdata_nested
```

<div style="margin-top:50px;">
##### This can be very useful for running, for example, a linear regression to try to predict the relationship between favorability and some other variable. And then mutating it to the nested table. I don't quite have the code down, but would love to see this extended :). 
###### favdata_nested <- favdata_nested |> mutate(regression=lm(sample_size ~ favorable, data=favdata_nested$data))
##### Note: a classification model may be more appropriate than a linear, to predict, say, the political candidate based on the polling information.
<div style="margin-top:50px;">