---
title: "Vignette Title"
author: "Karma Gyatso"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```



```{r}
library(tidyverse)
library(ggplot2)
```


## Loading data from github, which was downloaded from Kaggle. classifying the dataset type and printing all the names of the variables. 
```{r}
patient_records_test <- read.csv("https://raw.githubusercontent.com/karmaggyatso/CUNY_SPS/main/Github_data607/tidyverse_assignment/Train-1617360447408-1660719685476.csv")

```

```{r}
class(patient_records_test)
```

```{r}
names(patient_records_test)

```


## Research question 
**What is the main factor of patient readmission in the hospital? Are the treatment and age group the reason for readmission? What are the chances of readmission based on medical treatment and age group?**



## Cases 

**Each case represents individual patient who went to hospital for treatment. There are *66,587* observations**

```{r}
dim(patient_records_test)
```

## Resources
**Data source cited in MLA 9th edition**
 + Vutukuri, Girish. “Hospital_Administration.” Hospital_Administration | Kaggle, www.kaggle.com/datasets/girishvutukuri/hospital-administration/code?resource=download. Accessed 30 Oct. 2022.
 
 
## Analysis

**Data cleaning**
There is a data with medical_treatment as "?". Since we are not sure what sort of medical_treatement it is, we are omitting the records in the new variable. We're grouping the type of medical_treatement and counting how many patient has readmitted. 
```{r}
hospital_readmission_by_treatment <- patient_records_test %>% 
  filter(readmitted == 1) %>%
  filter(medical_specialty != "?") %>%
  group_by(medical_specialty) %>%
  count() %>% 
  arrange(desc(n))


hospital_readmission_by_treatment

```

A graphical representation of above code. 
```{r}
ggplot(hospital_readmission_by_treatment, aes(x = medical_specialty, y = n)) +
  geom_bar(stat = "identity") +
  theme(text = element_text(size = 6),element_line(size = 2)) +
  coord_flip() 
```

We also have to consider the age as an factor. So, I am grouping the age and counting total readmission. 
```{r}

hospital_readmission_by_age <- patient_records_test %>% 
  filter(readmitted == 1) %>%
  group_by(age) %>%
  count() %>% 
  arrange(desc(n))

hospital_readmission_by_age
```

A graphical representation of above code. 
```{r}
ggplot(hospital_readmission_by_age, aes(x = age, y = n)) +
  geom_bar(stat = "identity") + 
  labs(title= "Hospital readmission by age",
       x="age_group",y="total count")+
  coord_flip()
  
```

## Conclusion:
Those patients who had visited hospital for **internalMedicine** as a medical_treatment are most readmitted. Also the age group between 80-90 are also readmitted to the hospital. Further analysis can be done on age with most medical_treatment applied. 

I have used the tidyverse,dplyr fucntions such as piping fuction, filter and group_by function.  