---
title: "Data607_Tidyverse_Create_Assignment"
author: "Mahmud Hasan Al Raji"
date: "`r Sys.Date()`"
output: html_document
---

# The main task here is to Create an example by using one or more TidyVerse packages, and any dataset from fivethirtyeight.com or Kaggle, create a programming sample “vignette” that demonstrates how to use one or more of the capabilities of the selected TidyVerse package with the selected dataset.Here, I have selected a data set from kaggle.com and put that data set on my github. The data set reflects the different properties of two types of wine.





```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r }
library(tidyverse)
```

```{r }
df<-read.csv("https://raw.githubusercontent.com/Raji030/data607_tidyverse_create_assignment/main/winequalityN.csv")
glimpse(df)
```

# The dplyr package in tidyverse can be used to subset a data frame by subsetting rows using specific column value:
```{r }
read_wine_data<-df %>% filter(type=="red")
glimpse(read_wine_data)
```

# The dplyr package in tidyverse package can also be used to count the number of times a column value occurs:
```{r use_count_function}
count_wine_type<-df %>% count(type)
count_wine_type
```


# The ggplot2 package in tidyverse can be used to visualize relationship between variables of interest
```{r }
# Creating horizontal plot to visualize the count by wine types
ggplot(data=count_wine_type, aes(x=type, y=n)) +
  geom_bar(stat="identity") + 
    labs(
      x = "Wine Type",
      y = "Count",
      title = "The count of wine types"
    ) + 
    coord_flip()
```

# The purrr package is used to compute the summary of different variables
```{r }

df %>% split(.$type) %>% # from base R
  map(summary)
```


