---
title: "Data_607_TidyVerse_CREATE_Assignment_Vignette.rmd"
author: "Enid Roman"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data_607_TidyVerse_CREATE_Assignment.rmd}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

###**INTRODUCTION**

#### Which State Has The Worst Drivers?

###**DATA BACKGROUND**

#### This dataset, "bad-drivers" is from Fivethirtyeight. This data shows America’s bad drivers in all the states, involved in collisions. 

#### The data contains the following fields:

#### State
#### Number of drivers involved in fatal collisions per billion miles
#### Percentage Of Drivers Involved In Fatal Collisions Who Were Speeding
#### Percentage Of Drivers Involved In Fatal Collisions Who Were Alcohol-Impaired
#### Percentage Of Drivers Involved In Fatal Collisions Who Were Not Distracted
#### Percentage Of Drivers Involved In Fatal Collisions Who Had Not Been Involved In Any Previous Accidents
#### Car Insurance Premiums ($)
#### Losses incurred by insurance companies for collisions per insured driver ($)

###**RESOURCES**

#### Data is collected by the National Highway Traffic Safety Administration 2009 and 2012 (NHTSA) (https://www-fars.nhtsa.dot.gov/Main/index.aspx) and National Association of Insurance Commissioners 2010 & 2011(website was not provided).

#### Data is by Mona Chalabi from FiveThirtyEight and is available online here: https://github.com/fivethirtyeight/data/tree/master/bad-drivers For this project, the data was uploaded by calling the URL and using the read_table() function.

###**PACKAGES**

#### Package used is tidyverse. It includes readr, dplyr, tidyr, ggplot2, stringr, tibble, forcats and purr packages.

```{r}
#Upload tidyverse, ggplot2, and dplyr libraries.

library(tidyverse)
library(ggplot2)
library(dplyr)
library(infer)
```

###**ANALYSIS**

###**SECTION 1:LOAD THE DATA**

#### read_csv() function is from readr package, used for reading flat file data with comma separated values.

```{r}
# Load the dataset from fivethirtyeight.com and read the dataset with read.csv.

theUrl <- "https://raw.githubusercontent.com/fivethirtyeight/data/master/bad-drivers/bad-drivers.csv"

worst_drivers <- read.csv(file=theUrl, header=TRUE, sep=",")

head(worst_drivers)
```

###**SECTION 2: EXPLORATORY DATA ANALYSIS/CHECKING ASSUMPTION**
###**SECTION 3: SUMMARY STATISTICS**

#### glimpse() function is from tibble package, used to see column names, how many rows, how many columns, and type of variables in a data frame

```{r}
# Checking the columns name and how many rows and columns and variables type in that particular column. 

glimpse(worst_drivers)
```

#### colnames() is from dplyr package. Is used to obtain or set the names of the row and column of a matrix-like object, respectively.

```{r}
# Rename the columns for easier reading.

colnames(worst_drivers) <- c("STATE", 
                           "DRIVERS_INVOLVED", 
                           "PERC_DRIVERS_SPEED", 
                           "PERC_DRIVERS_ALCHO", 
                           "PERC_DRIVERS_NOT_DIST", 
                           "PERC_DRIVERS_NO_ACC", 
                           "INS_PREM", 
                           "LOSS_INSCOMP")

glimpse(worst_drivers)
```

### **DATA WRANGLING/VISUALIZATION USING DPLYR, TIDYR, AND GGPLOT2** 

#### Columns PERC_DRIVERS_SPEED, PERC_DRIVERS_ALCHO, PERC_DRIVERS_NOT_DIST, PERC_DRIVERS_NO_ACC are percentages of DRIVERS_INVOLVED. I will mutate new columns DRIVERS_SPEED, DRIVERS_ALCHO, DRIVERS_NOT_DIST, DRIVERS_NO_ACC by taking the given percentage of DRIVERS_INVOLVED column.

#### mutate() function is from dplyr package, adds new variables and preserves existing ones.

```{r}
# Create new column DRIVERS_SPEED which will be (DRIVERS_INVOLVED*PERC_DRIVERS_SPEED)/100

percent_worst_drivers <- worst_drivers %>% 
  mutate(DRIVERS_SPEED=(DRIVERS_INVOLVED*PERC_DRIVERS_SPEED)/100) %>% 
  mutate(DRIVERS_ALCHO=(DRIVERS_INVOLVED*PERC_DRIVERS_ALCHO)/100) %>% 
  mutate(DRIVERS_NOT_DIST=(DRIVERS_INVOLVED*PERC_DRIVERS_NOT_DIST)/100) %>% 
  mutate(DRIVERS_NO_ACC=(DRIVERS_INVOLVED*PERC_DRIVERS_NO_ACC)/100)

glimpse(percent_worst_drivers)
```

###**SECTION 4: VISUALIZING THE RESULTS**

#### Create a stacked bar plot using ggplot() method having states on X axis and DRIVERS_SPEED and DRIVERS_INVOLVED stacked together on Y axis. We use select() method to get required columns. The gather() method to make data long for DRIVERS_INVOLVED and DRIVERS_SPEED and finally use ggplot() to draw stacked bar plot.

#### select() function is from dplyr package that keeps only the variables we mention.
#### gather() function is from tidyr package that takes multiple columns and collapses into key-value pairs, duplicating all other columns as needed.
#### ggplot() All ggplot2 plots begin with a call to ggplot().

```{r}
percent_worst_drivers %>% 
  select(STATE, DRIVERS_INVOLVED, DRIVERS_SPEED) %>% 
  gather(type, value, DRIVERS_INVOLVED:DRIVERS_SPEED) %>% 
  ggplot(., aes(x = STATE,y = value, fill = type)) + 
  geom_bar(position = "stack", stat="identity") + 
  scale_fill_manual(values = c("red", "darkred")) + 
  ylab("Drivers involved in Fatal collision while Speeding") + 
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))
```

#### Below plot of for STATE vs INS_PREMIUM that used ggplot() method to creat a bar plot.

```{r}
percent_worst_drivers %>% 
  ggplot(., aes(x = STATE,y = INS_PREM)) + 
  geom_bar(position = "stack", stat="identity") + 
  ylab("Car Insurance Premium") + 
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))

```

###**CONCLUSION**

#### I discussed various packages and their functions to explore bad drivers dataset. For complete set details refer (https://www.tidyverse.org/).

#### **Resources**:

#### https://www.tidyverse.org/packages/
#### https://fivethirtyeight.com/features/which-state-has-the-worst-drivers/

###**SESSION INFORMATION**

```{r}
options(width = 100)
devtools::session_info()
```

