---
title: "Data607_Tidyverse_Create_Assignment"
author: "Mahmud Hasan Al Raji"
date: "`r Sys.Date()`"
output: html_document
---

# The main task here is to Create an example by using one or more TidyVerse packages, and any dataset from fivethirtyeight.com or Kaggle, create a programming sample “vignette” that demonstrates how to use one or more of the capabilities of the selected TidyVerse package with the selected dataset.Here, I have selected a data set from kaggle.com and put that data set on my github. The data set reflects the different properties of two types of wine.





```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r }
library(tidyverse)
```

```{r }
df<-read.csv("https://raw.githubusercontent.com/Raji030/data607_tidyverse_create_assignment/main/winequalityN.csv")
glimpse(df)
```

# The dplyr package in tidyverse can be used to subset a data frame by subsetting rows using specific column value:
```{r }
read_wine_data<-df %>% filter(type=="red")
glimpse(read_wine_data)
```

# The dplyr package in tidyverse package can also be used to count the number of times a column value occurs:
```{r use_count_function}
count_wine_type<-df %>% count(type)
count_wine_type
```


# The ggplot2 package in tidyverse can be used to visualize relationship between variables of interest
```{r }
# Creating horizontal plot to visualize the count by wine types
ggplot(data=count_wine_type, aes(x=type, y=n)) +
  geom_bar(stat="identity") + 
    labs(
      x = "Wine Type",
      y = "Count",
      title = "The count of wine types"
    ) + 
    coord_flip()
```

# The purrr package is used to compute the summary of different variables
```{r }

df %>% split(.$type) %>% # from base R
  map(summary)
```

### **Tidyverse Extended by Enid Roman**

### **Distribution of Single Variables**

### Wine Quality

#### Red wine sample is smaller. We know that number of observations for red and white are different in the dataset, but still we can see that for both colors it’s normal distribution with almost the same picks at 5 and 6 quality point.

```{r}
table(df$quality)
```

#### Here we use the function qplot() in ggplot2 part of Tidyverse Package is very similar to the basic plot() function from the R base package. It can be used to create and combine easily different types of plots.

```{r}
library(ggplot2)
qplot(quality, data = df, fill = type, binwidth = 1) +
  scale_x_continuous(breaks = seq(3,10,1), lim = c(3,10)) 
```
### Level of alcohol

#### Alcohol level distribution looks skewed. Again, red wine sample is smaller but it gives the same pattern of alcohol level distribution as white wines. Most frequently wines have 9.5%, mean is 10.49% of alcohol.

```{r}
summary(df$alcohol)
```


```{r}
qplot(alcohol, data = df, fill = type, binwidth = 0.5) +
    scale_x_continuous(breaks = seq(8,15,0.5), lim = c(8,15))
```
### Wine Density

#### Looking at ‘table’ summary we see that there are two outliers: 1.0103 and 1.03898. To see the distribution of density clearer I used log10 and limited the data. Now we can see that density distribution of white wine is bimodal and of red wine is normal.

```{r}
summary(df$density)
```

```{r}
qplot(density, data = df, fill = type, binwidth = 0.0002) +
    scale_x_log10(lim = c(min(df$density), 1.00370), 
                  breaks = seq(min(df$density), 1.00370, 0.002))
```

### **Distribution of Two and More Variables**

### Density of Quality by color

#### In out sample we have almost the same amount of red and white wines with quality ‘3’, ‘4’ and ‘9’, more red wines with quality ‘5’ and more white wines with quality “6”, “7” and “8”.

```{r}
qplot(quality, data = df, binwidth = 1, color = type, geom = "density") + 
    scale_x_continuous(breaks = seq(3, 9, 1))
```


