---
title: "Data607_Tidyverse_Create_Assignment"
author: "Mahmud Hasan Al Raji"
date: "`r Sys.Date()`"
output: html_document
---

# The main task here is to Create an example by using one or more TidyVerse packages, and any dataset from fivethirtyeight.com or Kaggle, create a programming sample “vignette” that demonstrates how to use one or more of the capabilities of the selected TidyVerse package with the selected dataset.Here, I have selected a data set from kaggle.com and put that data set on my github. The data set reflects the different properties of two types of wine.





```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r }
library(tidyverse)
```

```{r }
df<-read.csv("https://raw.githubusercontent.com/Raji030/data607_tidyverse_create_assignment/main/winequalityN.csv")
glimpse(df)
```

# The dplyr package in tidyverse can be used to subset a data frame by subsetting rows using specific column value:
```{r }
read_wine_data<-df %>% filter(type=="red")
glimpse(read_wine_data)
```

# The dplyr package in tidyverse package can also be used to count the number of times a column value occurs:
```{r use_count_function}
count_wine_type<-df %>% count(type)
count_wine_type
```


# The ggplot2 package in tidyverse can be used to visualize relationship between variables of interest
```{r }
# Creating horizontal plot to visualize the count by wine types
ggplot(data=count_wine_type, aes(x=type, y=n)) +
  geom_bar(stat="identity") + 
    labs(
      x = "Wine Type",
      y = "Count",
      title = "The count of wine types"
    ) + 
    coord_flip()
```

# The purrr package is used to compute the summary of different variables
```{r }

df %>% split(.$type) %>% # from base R
  map(summary)
```


## Extended Version Karma Gyatso

### Alcohol Level
Comparison between white and red wines interm of alcohol. White wines are largely produced between 9% and 10% alcohol

```{r}
summary(df$alcohol)
```

```{r}
ggplot(df, aes(x = alcohol, fill = type)) + 
  geom_bar(width = 0.5) + 
  facet_wrap(~type) +
  scale_x_continuous(breaks = seq(8,15,1), lim = c(8,15)) 
```


### Quality Level
The most common quality of wine produced is grade 6 type quality. May be its the average price and most sellable product. 

```{r}
ggplot(df, aes(x = quality, fill = type)) + 
  geom_bar() + 
  scale_x_continuous(breaks = seq(2,10,1), lim = c(2,10)) + 
  labs(
    title = "Wine Quality (0 - worst, 10 - best)"
  )
```


###Chloride level

Chloride distribution is left skewed. The mean is 0.05
```{r}
summary(df$chlorides)
```

```{r}
chloride_amount <- df |>
  select(type, chlorides)

chloride_amount

ggplot(chloride_amount, aes(x = chlorides, fill = type)) + 
  geom_bar() + 
  facet_wrap(~type) +
  labs(
    title = "Chloride amount in a wine"
  )
```
### Chloride vs Quality

Here we are checking what amount of chloride is used in quality rated wine. In average rating wine, which are between 4 - 6 has highest amount of chloride level for both white and red wines. Whereas in the high rated quality wine has less amount of chloride. 

```{r}
quality_vs_chlorides <- df |> 
  select(type, chlorides, quality)

max(quality_vs_chlorides$quality)

ggplot(quality_vs_chlorides, aes(x = chlorides, y = quality)) + 
  geom_point(aes(color = factor(type))) + 
  scale_x_continuous(breaks = seq(0.0,0.5,0.05), lim = c(0.0,0.5)) + 
  labs(
    title = "Chloride amount in a wine"
  )
```

### Density vs Quality

Here we are checking the density vs the quality of the wine. It seems that the most average wine, between 4 - 6, ratings are denser. However, we can also see that high quality wines are also denser but very few in quantity. 
```{r}
density_vs_quality <- df |> 
  select(type, quality, density)

max(density_vs_quality$density)

ggplot(density_vs_quality, aes(x = density, y = quality)) + 
  geom_point(aes(color = factor(type))) + 
  facet_wrap(~type) +
    scale_x_continuous(breaks = seq(0.99,1.01,0.01), lim = c(0.99,1.01)) + 
  labs(
    title = "Density amount vs quality in a wine"
  )

```




