---
title: "Extend Tidyverse_packages_Melissa_Bowman"
author: "Melissa Bowman"
date: "2022-11-09"
output: html_document
---

For this section of the TidyVerse project, an creating an extended TidyVerse example on a classmates existing file. Here, I used Khyati Naik’s tidyverse assignment.  The data used here was the House Rent Prediction Data on Kaggle link here:  https://www.kaggle.com/datasets/iamsouravbanerjee/house-rent-prediction-dataset. The TidyVerse package used for the extension was ggplot. 


Read libraries
``````{r load-packages, message=FALSE}
library(tidyverse)
```

Reading in the data frames using the **readr** package
```{r , message=FALSE}
data_path <- "https://raw.githubusercontent.com/Naik-Khyati/create_tv/main/data/House_Rent_Dataset.csv"

# read data
house_rent_data <- read_csv(data_path)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

```{r}
glimpse(house_rent_data)
str(house_rent_data)
```

Seeing if there is any relation between rent amount and size of apartment using ggplot. 
```{r}
ggplot(house_rent_data, aes(x = Rent, y = Size)) + 
  geom_point()
```
<br>
The outliers made it difficult to tell if there was any relationship therefore, I split the graph up by city, represented by different colors, and used the geom_smooth(method = lm) which will show regression lines with shaded confidence region.

```{r}
ggplot(house_rent_data, aes(x = Rent, y = Size, colour= City)) + 
  geom_point(shape=1) +
  geom_smooth(method=lm)
```
