---
title: 'tidyverse: using forcats to improve your ggplots'
author: "catlin"
date: "2/10/2021"
output: html_document
---

Note that although forcats is part of the "tidyverse", it is not automatically loaded when you run `library(tidyverse)`

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(forcats)
```

# Handy forcats functions for ggplot2

## Comparing followers of world religions

Source: https://en.wikipedia.org/wiki/List_of_religious_populations

I was looking for a simple dataset with count data for many items to demonstrate some basic forcats functions that are useful when creating plots.

```{r, warning=FALSE}
religions = read_csv("https://raw.githubusercontent.com/acatlin/data/master/religions.csv", 
                  col_names = FALSE) %>% 
    rename(religion = X1, followers = X2) %>% 
    mutate(millions_of_followers = followers/1000000.0) %>% 
  select(religion, millions_of_followers)

religions
```

## 1A: basic ggplot

Q: What are the most followed religions?
A: Use ggplot to compare religious populations

You can also embed plots, for example:

```{r}
religions %>% 
  ggplot(aes(x = religion, y = millions_of_followers)) + 
    geom_col(fill = "lightblue") + 
    labs(x = "religion", y = "millions of followers", 
           caption = "https://en.wikipedia.org/wiki/List_of_religious_populations") 
```
 
## 1B: How do I flip coordinates?

```{r}
religions %>% 
  ggplot(aes(x = religion, y = millions_of_followers)) + 
    geom_col(fill = "lightblue") + 
    labs(x = "religion", y = "millions of followers", 
           caption = "https://en.wikipedia.org/wiki/List_of_religious_populations") + coord_flip()
```

## 2A: How do I change sort order?
Revised by: Andy Catlin

Q: How do we change the chart to show the most followed religions first?
A: Use forcats::fct_reorder()

```{r}
library(forcats) 

ggplot(religions, aes(x = fct_reorder(religion, millions_of_followers), 
                      y = millions_of_followers)) + 
  geom_col(fill = "lightblue") + 
  labs(x = "religion", y = "millions of followers", 
      caption = "https://en.wikipedia.org/wiki/List_of_religious_populations") + 
  coord_flip()
```

## 2B: How do I combine less frequently used categories?

Q: How do we combine the less-followed religions into a single group?
A: Use forcats::fct_other()

```{r}
top5 = unlist(select(head(arrange(religions, desc(millions_of_followers)), 5), religion))

religions %>% 
  mutate(religion = fct_other(religion, keep = top5, other_level = "Other religions")) %>%     
  ggplot(aes(x = fct_reorder(religion, millions_of_followers), y = millions_of_followers)) +
    geom_col(fill = "lightblue") + 
    labs(x = "religion", y = "millions of followers", 
        caption = "https://en.wikipedia.org/wiki/List_of_religious_populations") + 
        coord_flip()
```

# Subsetting data with Dplyr Select

The data set used in this example is from Five Thirty Eight's 22-23 NBA predictions. To see the predictions go to:
https://projects.fivethirtyeight.com/2023-nba-predictions/
To download a copy of the data go to:
https://data.fivethirtyeight.com/#nba-forecasts

```{r}
nba_rating_data <- read_csv("https://raw.githubusercontent.com/adavidowitz100/FALL2022TIDYVERSE/main/nba-forecasts/nba_elo.csv")
head(nba_rating_data)
```
Any "tidy select" expression can be used to specify your interested subset of columns. Here are some examples:

## How do I subset to only specific columns of the dataframe by name using Dplyr?
If you only wanted to analyze the specific outcomes of certain games you would not need the full data set.
```{r}
nba_games <- nba_rating_data |> dplyr::select(date, season, team1, team2, score1, score2)
head(nba_games)
```

## How do I select all of the data except for specified columns by name?
For this analysis I do not need to keep note of playoff status.
```{r}
nba_rating_notplayoff <- nba_rating_data |> dplyr::select(!playoff)
head(nba_rating_data)
```

## How do I return all cloumns containing a criteria?
I want a dataset that only contains column names containing "elo"
```{r}
nba_elo <- nba_rating_data |> dplyr::select(contains("elo"))
head(nba_elo)
```

## How do I combine multiple subset criteria?
```{r}
nba_elo_notcarmelo <- nba_rating_data |> dplyr::select(contains("elo") & !contains("carm"))
head(nba_elo_notcarmelo)
```

Tidyverse Extend

```{r}
library(dplyr)
```

## How do I subset columns to incorporate contains but also factor in position of the string using Dplyr?
If you only want to find the statistics that the dataset has before the game was played or for one team
```{r}
nba_pre_stats <- nba_rating_data |> dplyr::select(ends_with('pre'))
head(nba_pre_stats)

nba_ends_one <- nba_rating_data |> dplyr::select(ends_with('1'))

head(nba_ends_one)

nba_starts_one <- nba_rating_data |> dplyr::select(starts_with('elo'))

head(nba_starts_one)
```


## How do I subset columns to incorporate more complex pattern matching using Dplyr?
If you don't want all of the elo statistics, but a specified subset based on string patterns
```{r}
nba_elo_spec <- nba_rating_data |> dplyr::select(matches('.*elo_'))
head(nba_elo_spec)

nba_elo_one <- nba_rating_data |> dplyr::select(matches('.*elo1.*'))
head(nba_elo_one)
```

## How do you subset columns for a specific group of strings in order?
If there are many fields and not a clear pattern to grab them you can pull specific adjacent columns
```{r}
nba_beg_col <- nba_rating_data |> dplyr::select(date:elo2_post)
head(nba_beg_col)
```



