---
title: "DATA 607: Data Acquisition and Management"
author: "Melvin Matanos, Fall 2022"
subtitle: "Week 10 - Tideyverse Create Assignment"
date: "10/27/2022"
output:
  html_document: default
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# The purrr package and map() function

purrr enhances R’s functional programming (FP) toolkit by providing a complete and consistent set of tools for working with functions and vectors. If you’ve never heard of FP before, the best place to start is the family of map() functions which allow you to replace many for loops with code that is both more succinct and easier to read. The best place to learn about the map() functions is the iteration chapter in R for data science.( https://purrr.tidyverse.org/)

The map functions transform their input by applying a function to each element of a list or atomic vector and returning an object of the same length as the input.(https://purrr.tidyverse.org/reference/map.html)

I will demosntrate the purrr::map() and purrr::pmap() function with the use of purrr library for this vignette.

The data set can be found in (https://github.com/CSSEGISandData/COVID-19).
(https://github.com/CSSEGISandData/COVID-19/blob/master/csse_covid_19_data/csse_covid_19_daily_reports_us/12-31-2021.csv) 

# Loadin the required libraries

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(stringr)
```


```{r}
covid_19_data <- readr::read_csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports_us/12-31-2021.csv")
```


```{r}
covid_19_data$Province_State <- replace_na(covid_19_data$Province_State, "")
covid_19_data <- covid_19_data %>% select(Province_State, Country_Region, Lat, Long_, Confirmed, Deaths) %>% arrange(desc(Confirmed)) 

head(covid_19_data)
```

# purrr:map()

This function will allow you to apply a function with a single argument to a vector.

 Let's us first create our function and for instance we will determined what is the total percentage of confirmed cases made up each location.Then finally we will add a new column of percent total.

```{r}
percent_total <- function(x) {
  return(x/sum(x) * 100)
}
```

 
```{r}
new_col <- covid_19_data %>% dplyr::select(Confirmed) %>% purrr::map(percent_total)
class(new_col)
new_col[[1]][1:10]
```

```{r}
covid_19_data <- covid_19_data %>% mutate("Percent_Total" = round(unlist(new_col),2))
covid_19_data
```

# purrr:pmap()

This function is a variation of map() that will allow you to apply a function with multiple arguments to a vector.

Let us create a function with a mutliple arguments. For example we will create a new column and concatenate the Providence_State column with the Country_Region and separate the output with comma.

```{r}
comma_function <- function(x, y) {
  if (x == "") {
    column_value <- y
  } else {
    column_value <- stringr::str_c(x, y, sep = ", ")
  }
  return(column_value)
}

``` 


```{r}
new_argument_list <- list(x = covid_19_data$Province_State,  y = covid_19_data$Country_Region )
covid_19_data <- covid_19_data %>% mutate("Location" = unlist(purrr::pmap(new_argument_list, comma_function)))
head(covid_19_data$Location, 5)
```


Display some rows on the final dataframe.

```{r}
covid_19_data <- covid_19_data %>% select(Location, Lat, Long_, Confirmed, Percent_Total, Deaths)
head(covid_19_data)
```








