---
title: "DATA607_Tidyverse_Project"
author: "Josh Iden, extended by Mahmud Hasan Al Raji"
date: "2022-10-26"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# TIDYVERSE PROJECT

This project uses data from [fivethirtyeight.com](https://projects.fivethirtyeight.com/republicans-trump-election-fraud/) to demonstrate tidyverse package capabilities.

```{r load-packages}
library(tidyverse)
```

## THE DATA

```{r}
url <- "https://raw.githubusercontent.com/fivethirtyeight/data/master/election-deniers/fivethirtyeight_election_deniers.csv"
df <- read.csv(url) 
head(df)
```

The *dplyr* package allows us to use pipes to transform the data:

```{r}
df <- df |>
  mutate(Stance = toupper(Stance))
```

It also let's us count the number of times a value occurs:

```{r}
stance_ct <- df |>
      count(Stance) |>
      arrange(desc(n))
stance_ct
```
The *ggplot* package allows us to visualize this without the transformation. 

```{r warning=FALSE}
ggplot(df) +
  geom_histogram(aes(x=Stance), stat="count") +
  coord_flip()
```
The *tidyr* package lets us pivot data:

```{r}
by_state <- df |> # using dplyr pipe
  group_by(State, Stance) |>
  summarise(count = n()) |>
  pivot_wider(names_from = Stance, values_from = count)
by_state
```

# Extended part summary

I will see the numbers of Republican nominee for different offices who took different stances on the 2020 election. I will also see the office wise nominee numbers who fully denied the election. Finally, I will find the incumbents by types for different stances and will also graphically represent the incumbent numbers by grouped barplot for each stance category.

# Subsetting the df dataframe with only Incumbent,Office, and Stance columns by using select function from dplyr

```{r }
df1<-df %>% select(Incumbent,Office,Stance)
head(df1)
```
# Finding nominee numbers for different offices who took different stances on the 2020 election by using group_by and summarize functions of dplyr 

```{r }
df2 <- df1 %>% 
  group_by(Office) %>% 
  summarize(count = n())
df2
```

# Finding the Republican nominee numbers by offices who fully denied the 2020 election by filtering the required data by using filter function of dplyr 

```{r}

df3<-df1 %>% filter(Stance=="FULLY DENIED") %>% 
group_by(Office) %>% summarize(count = n())
df3
```

# Finding Incumbent numbers for different stance cateory by using group_by annd summarize functions of dplyr package

```{r }
df3 <- df1 %>% 
  group_by(Stance,Incumbent) %>% 
  summarize(count = n())
df3
```

# Visualizing two types of incumbent numbers for each category of stance by grouped barplot using different functions of ggplot2 package

```{r }
ggplot(df3, aes(factor(Stance), count,fill=Incumbent)) + 
  geom_bar(stat="identity", position = "dodge") + theme(axis.text.x = element_text(angle = 90, size = 10))+
  scale_fill_brewer(palette = "Set1")
```




