---
title: "Dyplyr to Look at Crop Production & Climate Change"
author: "Sangeetha Sasikumar"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
##Importing Libraries
```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(tidyverse)
library(dplyr)
```

##Reading in the Data
In this vignette, I will try to manipulate data in R using Tidyverse's [`dplyr`](https://dplyr.tidyverse.org/) library. After glancing at the data I chose, I debated between ggplot2 and dyplyr, and just stuck with dyplyr. I chose the dataset titled "Crop Production & Climate Change" https://www.kaggle.com/datasets/thedevastator/the-relationship-between-crop-production-and-cli?resource=download

##What is dplyr?
The dplyr is an R-package to clean, manipulate and summarize data. It allows one to explore data and manipulate it fast. 

```{r}
data <-read.csv('/Users/Sangeetha/Downloads/crop_production.csv')
head(data)
class(data)
names(data)
str(data)
```

##Filter Method
The filter() method helps one to subset a data frame, while keeping all rows that satisfy the conditions we mention. Here, we are filtering for every row with location "USA" and making a new subset of the original dataframe called "usa". 
```{r}
usa<-data %>% 
  filter(LOCATION == "USA")
head(usa)
```

##Select Method
The select function helps us choose the fields of a dataframe we want to see. 
```{r}
head(data %>% 
  select(LOCATION, INDICATOR, SUBJECT, TIME), n=15)

```

##Mutate method
The mutate function makes new variables that are functions of existing variables. For example, here we made the new variable "latest_years" and wanted it to equal every year which is greater than 2010. The column displays true if the years are above 2010, if not false. 
```{r}

head(data %>% 
  select(LOCATION, INDICATOR, SUBJECT, TIME)%>% 
  mutate(latest_years =data$TIME>2010), n=25)

```

##Summarize method
The summarize method reduces multiple values down to a single summary. Here we take the mean of the Value column. 
```{r}
data%>%
  summarize(avg_val=mean(Value))

```

##Arrange method
The final method in Tidyverse's `dyplyr` package is "arrange". Arrange() puts the rows in the order we need them to be in. In the code below, the field "Value" was put in descending order. 
```{r}
data %>%
  group_by(INDICATOR) %>%
  arrange(desc(Value))

```

Here, the arrange function was applied on the subset 'usa' which was created above and arranged in descending order of the field "Value". 2025, 2024 and 2023 have the top 3 Values. 
```{r}
head(usa %>%
  arrange(desc(Value)), n=10)

```

Thank you for reading this document, I hope you will experiment with dplyr on your own datasets!



