---
title: "Tidyverse"
author: "Melissa Bowman"
date: "2022-10-28"
output: html_document
---

For the project TidyVerse, an example was created using mostly the dplyr package and the data from a FiveThirtyEight article titled *2022 Is Not Another ‘Year Of The Woman’*  (https://fivethirtyeight.com/features/2022-women-candidates-data/) and its data was located here: https://github.com/fivethirtyeight/data/tree/master/primary-project-2022

Load libraries.
```{r libraries}
library(tidyverse)
```


Reading in the data frames using the **readr** package
```{r pressure, read_csv}
url_republican <- "https://raw.githubusercontent.com/fivethirtyeight/data/master/primary-project-2022/rep_candidates.csv"
url_democratic <-"https://raw.githubusercontent.com/fivethirtyeight/data/master/primary-project-2022/dem_candidates.csv"


df_rep <- read.csv(url_republican)
df_dem <- read.csv(url_democratic) 
```

### Using the dplyr package for the following: <br>

Creating new columns 
```{r}
df_dem <- df_dem %>%
  mutate(political_party = "Democrat")

df_rep <- df_rep %>%
  mutate(political_party = "Republican")
```

Counting unique variable in each column 
```{r}
df_dem %>%
  filter("Gender" != "") %>%
  count(Gender, name  = "Count")

df_rep %>%
  filter("Gender" != "") %>%
  count(Gender, name  = "Count")
```

Combining data frame into one 
```{r}
df_bind <- bind_rows(df_rep, df_dem)
```

Creating a subset of specific columns into a new data frame
```{r}
df_subset_bind <- df_bind %>%
  select(Candidate, 
         Gender, 
         Race.1, 
         Race.2, 
         Race.3, 
         Incumbent, 
         State, 
         Office, 
         District, 
         political_party)
```


