---
title: "TidyVerse EXTEND Assignment"
author: "Sanielle Worrell"
date: "2022-11-12"
output: html_document
---

Assignment: Your task here is to Extend an Existing Example. Using one of your classmate's examples (as created above), extend his or her example with additional annotated code.

Analysis Performed: Classmate example used was Mahmud Hasan Al Raji's tidyverse assignment on wine types. I extended my classmate code on GGPLOT by showing how to calculate the average alcohol by wine type and plot this using ggplot.

Conclusion: White white has a higher percentage of alcohol compared to red wine.

## 

## 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# CLASSMATES PORTION

```{r }
library(tidyverse)
library(tidyr)
```

```{r }
df<-read.csv("https://raw.githubusercontent.com/Raji030/data607_tidyverse_create_assignment/main/winequalityN.csv")
glimpse(df)
```

The dplyr package in tidyverse can be used to subset a data frame by subsetting rows using specific column value:

```{r }
read_wine_data<-df %>% filter(type=="red")
glimpse(read_wine_data)
```

The dplyr package in tidyverse package can also be used to count the number of times a column value occurs:

```{r use_count_function}
count_wine_type<-df %>% count(type)
```

The ggplot2 package in tidyverse can be used to visualize relationship between variables of interest

```{r }
# Creating horizontal plot to visualize the count by wine types
ggplot(data=count_wine_type, aes(x=type, y=n)) +
  geom_bar(stat="identity") + 
    labs(
      x = "Wine Type",
      y = "Count",
      title = "The count of wine types"
    ) + 
    coord_flip()
```

# TIDYVERSE EXTEND PORTION

```{r}
## Calculating average alcohol of red vs white wine using dyplr:
 count_wine_type <- df %>%
  group_by(type)%>%
  summarise(count=n(), avg_alch = mean(alcohol))
 count_wine_type
 
```

```{r}
## Creating a ggplot to show the average alcohol by wine type
ggplot(count_wine_type, aes(x = avg_alch, y = count, fill = type)) + 
  geom_bar(stat = "identity")+scale_fill_manual(values=c("red", "pink"))+
  labs(title = "Average alcohol by wine type", x = "Avg Alcohol", y = "Count of wine type")+
  theme_bw()
```

\`\`\`
