---
title: 'Data607: Tidyverse Create'
author: 'jfor: Josh Forster'
date: "2022-10-26"
output: html_document
---


#### Displaying Tidyverse Functionality

The purpose of this vignette is to provide examples of different tidyverse functions to help with tidying and transforming data.

#### Load Libaries

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

#### Read in FiveThirtyEight NBA Elo data

The example data set in this case will be FiveThirtyEight's Elo statistics for all NBA seasons where data was available spanning back to 1946.

```{r nba_historical}
link = 'https://projects.fivethirtyeight.com/nba-model/nba_elo.csv'
bball_csv <- read_csv(link,show_col_types=FALSE)

```

#### Reviewing season game counts for entire dataset

Dplyr allows for chaining of commands using the pipe operator (%>% or |>) to list multiple processes in a logical flow.

```{r review_summary}
bball_csv |>
    dplyr::group_by (season) |>
    dplyr::summarize(cnt=n()) |>
    dplyr::arrange(desc(season))
```
#### 1A: Filtering dataframes to review past 3 complete seasons

Given that the current season only just begun in October 2022, we will look at the prior 3 seasons for these examples

```{r}
last_three_seasons <- bball_csv |>
    dplyr::filter(season %in% c(2020,2021,2022))
#alternatively
last_three_seasons_v2 <- bball_csv |>
    dplyr::filter(season>=2020 & season<2023)

dim(last_three_seasons) == dim(last_three_seasons_v2)
```

The filter command allows for a number of different ways to slice the same data giving considerable flexibility to the end user. The dimensions are confirmed to be the same for either case 


#### 1B: Select only the columns that are useful for the specified analysis

Similar to dplyr's filter, the select function provides many alternatives to accomplish the same task depending on user preference. It allows for custom subset of columns to keep within a dataframe

```{r}
last_three_seasons |> select(date,season,team1,team2,elo1_pre,elo2_pre,elo1_post,elo2_post,score1,score2)
```
#### 1C: Aggregate data for review of comparative statistics

One question to potentially answer is which teams outperformed their Elo ratings and increased their total Elo for the season. By using group_by and summarize, we can choose specific variables to group on before we calculate minimum and maximum elo ratings for each team1 and season.

```{r}
sum_stats <- last_three_seasons |>
    group_by(team1,season) |>
    summarize(min_elo = min(elo1_pre),
              max_elo = max(elo1_post),
              season_diff = max_elo - min_elo) |>
    arrange(-season_diff)

```

We can sort the data within the dataframe to review the teams with the best improvement, but it doesn't necessarily give us a full picture except providing the top values.

#### 2: Build a ggplot scatterplot to review the league wide changes in these metrics

Building a base plot using ggplot, the geom_point is added as a layer to show a point within the scatterplot based on the x and y coordinating provided in the initial aesthetic which in this case is the minimum and maximum elo rating of a team. The specific aesthetics for the points themselves are provided within the geom_point instantiation which in this case adjusted the shape and size of the points to be driven off specific numeric variables in the dat.

```{r}
sum_stats %>%
    ggplot(mapping=aes(min_elo,max_elo)) +
    geom_point(aes(shape=as.character(season),size=season_diff)) +
    geom_text(aes(label=team1),vjust=-1.5, size=2) +
    labs(title='Elo Ratings over 3 prior NBA Seasons',caption='538 Data Set',size='Season Elo Difference',shape='Season')
```

Based on the initial head summary from the earlier chunk and this additional plot it is much more apparent that teams with lower initial expectations are typically the one with the biggest variability with in-season Elo ratings. This makes a bit of sense given that the model was somewhat bearish on these teams and they outperformed expectations during the season that were not initially accounted at their lowest point. There is a clear linear trend in the data overall and in general teams are improving their ratings over the full season.

*Filter* is the same function as the subset function, and is used to filter data by specific conditions. Below is an example of extracting only New York from the data.

```{r}
df %>% filter(State == "New York") 
```

*Arrange* is a function that sorts data according to a specific order. Unlike the order function, which only tells the sort order, it shows sorted data.

```{r}
df %>% arrange(Candidate)
```

*select* is a function that selects specific variables from data and provides useful functions that are not in basic R.
```{r}
df %>% select(Candidate, Incumbent, State)
```