---
title: "TIDYVERSE VIGNETTE"
author: "Keeno Glanville"
date: "2022-11-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggplot2)
```
# Tidyverse Vignette

## Load kaggle set Data
Github:https://raw.githubusercontent.com/kglan/FALL2022TIDYVERSE/main/UCI_Credit_Card.csv
kaggle: https://www.kaggle.com/datasets/uciml/default-of-credit-card-clients-dataset

## Information on the Dataset

* SEX: Gender (1=male, 2=female)
* EDUCATION: (1=graduate school, 2=university, 3=high school, 4=others, 5=unknown, 6=unknown)
* MARRIAGE: Marital status (1=married, 2=single, 3=others)
* AGE: Age in years
```{r}
ccd <- read.csv("https://raw.githubusercontent.com/kglan/FALL2022TIDYVERSE/main/UCI_Credit_Card.csv")
names(ccd)
```


## Research question 

**What are the strongest contributing factors(nominal) towards client defaulting on their payments? Are there any trends?**



## Cases 
 
**Each case represents individuals with credit card debt and their current status with defaulting on their payment. There are *30,000* observations**

```{r}
dim(ccd)
head(ccd)
ccd_n <- ccd%>%
  select(1,3,4,5,6,25)
head(ccd_n)
ccd1<- ccd_n%>%
  filter(default.payment.next.month == 1)
ccd0<- ccd_n%>%
  filter(default.payment.next.month == 1)
```


## Analysis

** How can we better understand this dataset and its correlations with nominal features?

```{r}
pairs(data = ccd, ~ ID + SEX + EDUCATION+ MARRIAGE + AGE+ default.payment.next.month)
```


## We now want to see how the Data correlates with default in the next month payment
```{r}

ggplot(ccd_n, aes(x = AGE, y = default.payment.next.month)) +
  geom_bar(stat = "identity") +
  theme(text = element_text(size = 6),element_line(size = 2))

ggplot(ccd_n, aes(x = SEX, y = default.payment.next.month)) +
  geom_bar(stat = "identity") +
  theme(text = element_text(size = 6),element_line(size = 2))

ggplot(ccd_n, aes(x = EDUCATION, y = default.payment.next.month)) +
  geom_bar(stat = "identity") +
  theme(text = element_text(size = 6),element_line(size = 2))

ggplot(ccd_n, aes(x = MARRIAGE, y = default.payment.next.month)) +
  geom_bar(stat = "identity") +
  theme(text = element_text(size = 6),element_line(size = 2))


ggplot(ccd_n, aes(x = AGE))+
  geom_histogram()
ggplot(ccd_n, aes(x = EDUCATION))+
  geom_histogram()



```

## We Notice these factors have high instances of bankruptcy:
* age ranges between 20-30 are more likely to default on their 
next payment.
* Individuals who went to university are most likely to default (followed by graduates)
* Females are more at fault
* With married and single individuals not having too much difference



## Conclusion:
Utilizing the ggplot2 function we were able to determine that the most at risk for defaults
on their next credit card payments could be women between the ages of 20-30 and have been through
higher education.
