---
title: "TidyVerse CREATE assignment"  
author: "Seung Min Song"
date: "2022-10-30"
output:
  html_document:
    df_print: paged
  latex_engine: xelatex
  pdf_document: null
fontsize: 10pt
linestretch: 1.5
---

GitHub: <https://github.com/seung-m1nsong/607>\
rpubs: <https://rpubs.com/seungm1nsong/963598>
  
```{r setup, include=FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(magrittr)
```
\
The tidyverse is quickly replacing the original R syntax with the advantage of being able to write intuitive code. The tidyverse package is a package that helps you install and load R packages that follow the tidy data paradigm at once. tidyverse is a package that installs and manages core packages belonging to the tidy package ecosystem, such as dplyr tidyr ggplot2, at once.
\

## Intuitive code: %>% operator
\
If I had to pick just one of the most important things in the tidyverse ecosystem, I would choose the %>% operator from the magrittr package. %>% can be entered with the shortcut Ctrl + Shift + M (OS X: Cmd + Shift + M) in Rstudio, and by using this, you can code in a stream of consciousness and write intuitive code. Let's find out the advantages of %>% through the head function.
```{r}
url <- "https://raw.githubusercontent.com/seung-m1nsong/607/main/tidyverse/car_purchasing.csv"
dfdata <- read.csv(url) 
```
```{r}
## head(dfdata)
dfdata %>% head
```

```{r}
dfdata %>% head(n = 10)
```
This command is much more intuitive than the existing head(subset(a, State == "NY")) as it shows the head by selecting only *NY* from *dfdata*.
```{r}
## head(subset(a, gender == "1"))
dfdata %>% subset(gender == "1") %>% head
```

## Data Cleanup: dplyr
\
dplyr provides a set of functions to manipulate data effectively. Of these, group_by and summarize provide a differentiated value from the existing R syntax by easily showing summary statistics for each group. 

### filter
\
filter is the same function as the subset function, and is used to filter data by specific conditions. Below is an example of extracting only men from the data.

```{r}
dfdata %>% filter(gender == "1") 
```

In filter, you can use AND conditions with , in addition to &, so readability is good. You can also select a specific range of a continuous variable by using the between function, which is also more intuitive than using the existing &. Let's look at an example of filtering between 50 and 60 years old.

```{r}
## Age between 50 and 60.
dfdata %>% filter(age >= 50 & age <= 60)
```

### arrange: sort
\
arrange is a function that sorts data according to a specific order. Unlike the order function, which only tells the sort order, it shows sorted data.
```{r}
## dfdata[order(dfdata$Age), ]
dfdata %>% arrange(age)
```
If there are two or more sort conditions, you can write them together with , and use the desc command to sort in descending order. Below is an example of sorting in ascending order on Age and descending order on gender.
```{r}
## dfdata[order(dfdata$age, -dfdata$gender), ]
dfdata %>% arrange(age, desc(gender))
```

### mutate: create a variable
\
mutate is a function that creates a new variable. Let's create Old and Overweight variables that mean old age and obesity from Age and BMI variables.
```{r}
## dfdata$old <- as.integer(a$age >= 65); dfdata$middleclass <- as.integer(dfdata$annual_Salary >= 80000)
dfdata %>% mutate(Old = as.integer(age >= 50),
             middleclass = as.integer(annual_Salary >= 80000)
             )
```
To show only new variables, use transmute instead of mutate.
```{r}
dfdata %>% transmute(Old = as.integer(age >= 50),
             middleclass = as.integer(annual_Salary >= 80000)
             )
``` 

### group_by and summarize
\
By using group_by and summarize, you can divide groups as desired and obtain summary statistics for each group. In basic R, the aggregate function performs the same function.
```{r}
dfdata %>% 
  group_by(age, gender) %>% 
  summarize(count = n(),
            meanSalary = mean(annual_Salary))
```
To insert a string such as "age" in group_by, use the group_by_ function with an underscore (_).

## Conclusion
\
So far, we have seen how to manipulate data using several packages in the tidyverse ecosystem. As I said earlier, the most important thing in this ecosystem is to perform coding according to the stream of consciousness using the %>% operator, and if you apply the rest of the contents one by one, at some point you will find yourself unable to live without the tidyverse. 



