
---
title: "TIDYVERSE VIGNETTE"
author: "Keeno Glanville"
date: "2022-11-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggplot2)
```
# Tidyverse Vignette

## Load kaggle set Data
Github:https://raw.githubusercontent.com/kglan/FALL2022TIDYVERSE/main/UCI_Credit_Card.csv
kaggle: https://www.kaggle.com/datasets/uciml/default-of-credit-card-clients-dataset

## Information on the Dataset

* SEX: Gender (1=male, 2=female)
* EDUCATION: (1=graduate school, 2=university, 3=high school, 4=others, 5=unknown, 6=unknown)
* MARRIAGE: Marital status (1=married, 2=single, 3=others)
* AGE: Age in years
```{r}
ccd <- read.csv("https://raw.githubusercontent.com/kglan/FALL2022TIDYVERSE/main/UCI_Credit_Card.csv")
names(ccd)
```


## Research question 

**What are the strongest contributing factors(nominal) towards client defaulting on their payments? Are there any trends?**



## Cases 
 
**Each case represents individuals with credit card debt and their current status with defaulting on their payment. There are *30,000* observations**

```{r}
dim(ccd)
head(ccd)
ccd_n <- ccd%>%
  select(1,3,4,5,6,25)
head(ccd_n)
ccd1<- ccd_n%>%
  filter(default.payment.next.month == 1)
ccd0<- ccd_n%>%
  filter(default.payment.next.month == 1)
```


## Analysis

** How can we better understand this dataset and its correlations with nominal features?

```{r}
pairs(data = ccd, ~ ID + SEX + EDUCATION+ MARRIAGE + AGE+ default.payment.next.month)
```


## We now want to see how the Data correlates with default in the next month payment
```{r}

ggplot(ccd_n, aes(x = AGE, y = default.payment.next.month)) +
  geom_bar(stat = "identity") +
  theme(text = element_text(size = 6),element_line(size = 2))

ggplot(ccd_n, aes(x = SEX, y = default.payment.next.month)) +
  geom_bar(stat = "identity") +
  theme(text = element_text(size = 6),element_line(size = 2))

ggplot(ccd_n, aes(x = EDUCATION, y = default.payment.next.month)) +
  geom_bar(stat = "identity") +
  theme(text = element_text(size = 6),element_line(size = 2))

ggplot(ccd_n, aes(x = MARRIAGE, y = default.payment.next.month)) +
  geom_bar(stat = "identity") +
  theme(text = element_text(size = 6),element_line(size = 2))


ggplot(ccd_n, aes(x = AGE))+
  geom_histogram()
ggplot(ccd_n, aes(x = EDUCATION))+
  geom_histogram()



```

## We Notice these factors have high instances of bankruptcy:
* age ranges between 20-30 are more likely to default on their 
next payment.
* Individuals who went to university are most likely to default (followed by graduates)
* Females are more at fault
* With married and single individuals not having too much difference



## Conclusion:
Utilizing the ggplot2 function we were able to determine that the most at risk for defaults
on their next credit card payments could be women between the ages of 20-30 and have been through
higher education.
=======
---
title: "TIDYVERSE VIGNETTE"
author: "Keeno Glanville"
date: "2022-11-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggplot2)
```
# Tidyverse Vignette

## Load kaggle set Data
Github:https://raw.githubusercontent.com/kglan/FALL2022TIDYVERSE/main/UCI_Credit_Card.csv
kaggle: https://www.kaggle.com/datasets/uciml/default-of-credit-card-clients-dataset

## Information on the Dataset

* SEX: Gender (1=male, 2=female)
* EDUCATION: (1=graduate school, 2=university, 3=high school, 4=others, 5=unknown, 6=unknown)
* MARRIAGE: Marital status (1=married, 2=single, 3=others)
* AGE: Age in years
```{r}
ccd <- read.csv("https://raw.githubusercontent.com/kglan/FALL2022TIDYVERSE/main/UCI_Credit_Card.csv")
names(ccd)
```


## Research question 

**What are the strongest contributing factors(nominal) towards client defaulting on their payments? Are there any trends?**



## Cases 
 
**Each case represents individuals with credit card debt and their current status with defaulting on their payment. There are *30,000* observations**

```{r}
dim(ccd)
head(ccd)
ccd_n <- ccd%>%
  select(1,3,4,5,6,25)
head(ccd_n)
ccd1<- ccd_n%>%
  filter(default.payment.next.month == 1)
ccd0<- ccd_n%>%
  filter(default.payment.next.month == 1)
```


## Analysis

** How can we better understand this dataset and its correlations with nominal features?

```{r}
pairs(data = ccd, ~ ID + SEX + EDUCATION+ MARRIAGE + AGE+ default.payment.next.month)
```


## We now want to see how the Data correlates with default in the next month payment
```{r}

ggplot(ccd_n, aes(x = AGE, y = default.payment.next.month)) +
  geom_bar(stat = "identity") +
  theme(text = element_text(size = 6),element_line(size = 2))

ggplot(ccd_n, aes(x = SEX, y = default.payment.next.month)) +
  geom_bar(stat = "identity") +
  theme(text = element_text(size = 6),element_line(size = 2))

ggplot(ccd_n, aes(x = EDUCATION, y = default.payment.next.month)) +
  geom_bar(stat = "identity") +
  theme(text = element_text(size = 6),element_line(size = 2))

ggplot(ccd_n, aes(x = MARRIAGE, y = default.payment.next.month)) +
  geom_bar(stat = "identity") +
  theme(text = element_text(size = 6),element_line(size = 2))


ggplot(ccd_n, aes(x = AGE))+
  geom_histogram()
ggplot(ccd_n, aes(x = EDUCATION))+
  geom_histogram()



```

## We Notice these factors have high instances of bankruptcy:
* age ranges between 20-30 are more likely to default on their 
next payment.
* Individuals who went to university are most likely to default (followed by graduates)
* Females are more at fault
* With married and single individuals not having too much difference



## Conclusion:
Utilizing the ggplot2 function we were able to determine that the most at risk for defaults
on their next credit card payments could be women between the ages of 20-30 and have been through
higher education.

## Mathew Katz Extension:
```{r}
head(ccd)
```

```{r}
# Drawing a list of the variables that need to be changed
categorical = c('SEX','EDUCATION', 'MARRIAGE', 'default.payment.next.month',
               colnames(ccd)[grep("PAY_\\d", colnames(ccd))])

# Changing variables type
ccd[,categorical] = lapply(ccd[,categorical], function(var) as.factor(var))
#str(ccd)
summary(ccd)
```
```{r}
# Creating a table of quantitative variables
quantitative = ccd[, !colnames(ccd) %in% categorical]
df = cbind.data.frame(quantitative, target = ccd$default.payment.next.month)

# Creating a table of categorical variables
qualitative = ccd[,categorical]
```

```{r}
quantitative
```
```{r}
qualitative
```

```{r}
#Renaming column PAY_0 to PAY_1 and default.payment.next.month as DEFAULT

colnames(ccd)[colnames(ccd) == "PAY_0"] = "PAY_1"

colnames(ccd)[colnames((ccd)) == "default.payment.next.month"] = "DEFAULT"
```

```{r}
#Convert int variables to factors
ccd$DEFAULT = as.factor(ccd$DEFAULT)
```

```{r}
ccd$SEX = as.factor(ccd$SEX)
```

```{r}
table(ccd$DEFAULT)
```

```{r}
#Add one feature GENDER
ccd$GENDER = ifelse(ccd$SEX == 1, "Male", "Female")
```

```{r}
# Bar Graph for gender
ggplot(data = ccd, mapping = aes(x = GENDER, fill = DEFAULT)) +
  geom_bar() +
  ggtitle("Gender") +
  stat_count(aes(label = ..count..), geom = "label")
```
```{r}
#Merging 0, 5 and 6 to 4(others)
ccd$EDUCATION = ifelse(ccd$EDUCATION == 0 |ccd$EDUCATION == 5 | ccd$EDUCATION == 6,
       4, ccd$EDUCATION)
```

```{r}
#Converting EDUCATION to categorical variable
ccd$EDUCATION = as.factor(ccd$EDUCATION)
```

```{r}
ggplot(data = ccd, mapping = aes(x = EDUCATION, fill = DEFAULT)) +
  geom_bar() +
  ggtitle("EDUCATION") +
  stat_count(aes(label = ..count..), geom = "label")
```












=======
---
title: "TIDYVERSE VIGNETTE"
author: "Keeno Glanville"
date: "2022-11-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggplot2)
```
# Tidyverse Vignette

## Load kaggle set Data
Github:https://raw.githubusercontent.com/kglan/FALL2022TIDYVERSE/main/UCI_Credit_Card.csv
kaggle: https://www.kaggle.com/datasets/uciml/default-of-credit-card-clients-dataset

## Information on the Dataset

* SEX: Gender (1=male, 2=female)
* EDUCATION: (1=graduate school, 2=university, 3=high school, 4=others, 5=unknown, 6=unknown)
* MARRIAGE: Marital status (1=married, 2=single, 3=others)
* AGE: Age in years
```{r}
ccd <- read.csv("https://raw.githubusercontent.com/kglan/FALL2022TIDYVERSE/main/UCI_Credit_Card.csv")
names(ccd)
```


## Research question 

**What are the strongest contributing factors(nominal) towards client defaulting on their payments? Are there any trends?**



## Cases 
 
**Each case represents individuals with credit card debt and their current status with defaulting on their payment. There are *30,000* observations**

```{r}
dim(ccd)
head(ccd)
ccd_n <- ccd%>%
  select(1,3,4,5,6,25)
head(ccd_n)
ccd1<- ccd_n%>%
  filter(default.payment.next.month == 1)
ccd0<- ccd_n%>%
  filter(default.payment.next.month == 1)
```


## Analysis

** How can we better understand this dataset and its correlations with nominal features?

```{r}
pairs(data = ccd, ~ ID + SEX + EDUCATION+ MARRIAGE + AGE+ default.payment.next.month)
```


## We now want to see how the Data correlates with default in the next month payment
```{r}

ggplot(ccd_n, aes(x = AGE, y = default.payment.next.month)) +
  geom_bar(stat = "identity") +
  theme(text = element_text(size = 6),element_line(size = 2))

ggplot(ccd_n, aes(x = SEX, y = default.payment.next.month)) +
  geom_bar(stat = "identity") +
  theme(text = element_text(size = 6),element_line(size = 2))

ggplot(ccd_n, aes(x = EDUCATION, y = default.payment.next.month)) +
  geom_bar(stat = "identity") +
  theme(text = element_text(size = 6),element_line(size = 2))

ggplot(ccd_n, aes(x = MARRIAGE, y = default.payment.next.month)) +
  geom_bar(stat = "identity") +
  theme(text = element_text(size = 6),element_line(size = 2))


ggplot(ccd_n, aes(x = AGE))+
  geom_histogram()
ggplot(ccd_n, aes(x = EDUCATION))+
  geom_histogram()



```

## We Notice these factors have high instances of bankruptcy:
* age ranges between 20-30 are more likely to default on their 
next payment.
* Individuals who went to university are most likely to default (followed by graduates)
* Females are more at fault
* With married and single individuals not having too much difference



## Conclusion:
Utilizing the ggplot2 function we were able to determine that the most at risk for defaults
on their next credit card payments could be women between the ages of 20-30 and have been through
higher education.


